Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-03-17 VegOzone.tex
First draft of dos3e results

The methods described above rely on accurate representation of plant physiological response to climate variables (i.e. the species or plant functional type specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates. Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable. In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as reference.
It is important to understand the implications of using these more generic parameterizations for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization. we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change. This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-geochemcial cycling. This might be one key to solve long-standing issues of coupled earth system models (ESM) which generally lack PFTs representing subarctic conditions in their land surface model.\\

@@

\begin{figure}[t]
  \includegraphics[width=12cm]{fig08}
  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standart deviation for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for precip.). (a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$. The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details). The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July. Ozone was significantly elevated in May. The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August. No significant elevation in ozone was found.}
  \label{fig:anomalies_svanvik}
\end{figure}

@@

\section{DO3SE modeling}
\label{sec:do3se}
To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic prameterizations for common plant species. We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used method for deriving the bespoke parameters. Then we will present and discuss the modeling results.

@@

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence during a short, subarctic growing season and hence their carbon sequestration. Because the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegation is subject to ongoing climate change and acclimations may be changing. Resulting reponse functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

We define two different hypothetical temperature acclimations: \emph{Boreal} and \emph{cold}. For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000). For perennial grassland the resulting response functions are depicted together with the PDF of temperatur in Fig.~\ref{fig:f_temp_grassland}a).

@@

For an acclimation with respect to light conditions, we hypothize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species. We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$. The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for downy birch and Norway spruce refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To quantify the resulting bespoke parameterizations, we propose the following metric. For a generic GS (May--August), we compute climatological average and standart deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time). We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency. Our arbitrary optimization target is $60\,\unit{\%}$. Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions. For simplicity, we neglect the dependency on SWP in this regard. From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike. Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergance for perannial grassland is substantial.

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standart deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland. For simplicity, $f_\mathrm{SW}$ is neglected in this calculation. }
  \label{fig:javis_func_opt_mean}
\end{figure}

Another important factor for ozone exposure metrices is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincident with the first snow-free day at the meterological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.

@@

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variaty of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown comprehensively in Fig.~\ref{fig:pody_rel}. For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce), and $10\,\unit{\%}$ (perennial grassland, above ground) reduction in biomass \citep{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. Different years are color coded as previously. Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles. The annotation $\pm$ refers to bespoke light response function threshold ($+$ increased, $-$ decreased), while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing temperature acclimation $\rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\rightarrow$ increased uptake,
\item drought conditions (SWP) $\rightarrow$ reduced uptake,
\item lower/higher light threshold $\rightarrow$ incresed/reduced uptake,
\item bespoke light threshold $\rightarrow$ asymmetric response in POD,
\item temperature acclimation $\rightarrow$ amplified effect of light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species. Drought effects (SWP) do only matter if a bespoke GS is not taken into account.

In more detail for downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest range of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2019. Without bespoke GS 2018 stands out as a higher ozone damage risk year compared to 2019. However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. SWP induced reduction in ozone uptake is only effectiv in 2018 and without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased. Norway spruce displayes a medium sensitivity to the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarily large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perannial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhoved (2018, 2019). (a) Downy birch; (b) Norway spruce; (c) Perennial grassland. The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

The estimated biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017} is shown in Table~\ref{tab:biomass_reduction}. {\bf TODO: Lisa? Add paragraph about biomass reduction to introduction?}

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} due to \chem{O_3} following \citet{ICP:MappingManual2017}.}
  \label{tab:biomass_reduction}
\begin{tabular}{lccrr}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & 11.99 & 2.34 & 8.06 & 10.51$^*$\\
2019 & 10.07 & 2.26 & 6.43 & 7.90$^*$\\
clim. & 13.81 & 2.58 & 6.67 & 8.29$^*$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

>>>>

The methods described above rely on accurate representation of plant physiological response to climate variables (i.e. the species or plant functional type specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates. Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable. In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as reference.
It is important to understand the implications of using these more generic parameterization for more extreme climate regions, here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization. we also explore the role \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change. This allows both an understanding of how the threat from ozone may change over time but also which are the key aspects of plant physiology that might determine risk (and hence which parameters might require special attention to get right) both for pollution impact assessment, but also more generally for describing gas exchange response to changing climates that will influence bio-chemical cycling. This might be one key to solve long-standing issues of coupled earth system models (ESM) which generally lack PFTs representing subarctic conditions in their land surface model.\\

@@

\begin{figure}[t]
  \includegraphics[width=12cm]{fig08}
  \caption{2018/19 anomalies of key environmental variables at Svanhovd displayed as percentages of days deviating from climatological mean by means of standard deviation for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The hatched area between the dashed lines indicates statistical significance under the assumption of a normal distribution ($15.9\,\unit{\%}$, $30.8\,\unit{\%}$ for Precip.). (a) Temperature, days warmer/colder than $\pm 1\,\sigma$; (b) precipitation, days wetter/drier than $\pm \frac{1}{2}\,\sigma$; (c) global irradiance, days lighter/darker than $\pm 1\,\sigma$; (d) ozone, days higher/lower than $\pm 1\,\sigma$. The July 2018 ozone data corrected for data acquisition problems is indicated with the star (refer Section~\ref{subsec:ozone_reco} for details). The growing season of 2018 was significantly warmer, brighter, and slightly drier than average, especially in May and July. Ozone was significantly elevated in May. The growing season of 2019 was fairly average but significantly warmer than average in April and colder in July/August. No significant elevation in ozone was found.}
  \label{fig:anomalies_svanvik}
\end{figure}

@@

\section{DO3SE modeling}
\label{sec:do3se}
To investigate which of the environmental factors lead to the observed ozone damage on clover and tobacco in 2018 in contrast to 2019, when no such damage was found on the clovers but only on the sensitive tobacco cultivar, we model ozone uptake of different natural and semi-natural vegetation using the $\mathrm{DO_3SE}$ model and develop bespoke subarctic parameterizations for common plant species. We first give an account of the $\mathrm{DO_3SE}$ model, input data, and the used method for deriving the bespoke parameters. Then we will present and discuss the modeling results.

@@

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence during a short, subarctic growing season and hence their carbon sequestration. Because the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and acclimation may be changing. Resulting response functions for Norway spruce and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

We define two different hypothetical temperature acclimation: \emph{Boreal} and \emph{cold}. For either, we assume that the hypothetical species are more tolerant to cold conditions and shift $f_\mathrm{T}$ towards colder temperatures accordingly. We construct \emph{cold} as representative for a species that is more tolerant to cold temperatures, but slightly less efficient at warm temperatures compared to MM. In the same way, \emph{boreal} refers to a hypothetical species which is very tolerant to cold but is sensitive to high temperatures, and most efficient at cool temperatures ($T_\mathrm{opt}$ close to the climatological mean temperature of 1992--2000). For perennial grassland the resulting response functions are depicted together with the PDF of temperature in Fig.~\ref{fig:f_temp_grassland}a).

@@

For an acclimation with respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species. We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$. The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for downy birch and Norway spruce refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To quantify the resulting bespoke parameterizations, we propose the following metric. For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time). We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency. Our arbitrary optimization target is $60\,\unit{\%}$. Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions. For simplicity, we neglect the dependency on SWP in this regard. From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike. Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial.

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning ($5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland. For simplicity, $f_\mathrm{SW}$ is neglected in this calculation. }
  \label{fig:javis_func_opt_mean}
\end{figure}

Another important factor for ozone exposure metrics is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and shall be used for all vegetation alike. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincident with the first snow-free day at the meteorological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by. From the latitude model typically used for localizing $\mathrm{DO_3SE}$, we get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ indicating an about $2\,\unit{months}$ longer growing season. Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.

@@

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown comprehensively in Fig.~\ref{fig:pody_rel}. For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce), and $10\,\unit{\%}$ (perennial grassland, above ground) reduction in biomass \citep{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. Different years are color coded as previously. Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles. The annotation $\pm$ refers to bespoke light response function threshold ($+$ increased, $-$ decreased), while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

The results depicted in Fig.~\ref{fig:pody_rel} allow for a comprehensive assessment of a large variety of systematic uncertainties in the exposure-based risk metric for ozone induced damage in subarctic vegetation. In summary, we identify the following main effects on the model predicted ozone uptake with respect to MM default parameterization:
\begin{itemize}
  \itemsep0pt
\item Increasing temperature acclimation $\Rightarrow$ increased uptake regardless of individual year,
\item usage of proper growing season $\Rightarrow$ reduced uptake with respect to the above,
\item an earlier assumed onset of growing season $\Rightarrow$ increased uptake,
\item drought conditions (SWP) $\Rightarrow$ reduced uptake,
\item lower/higher light threshold $\Rightarrow$ increased/reduced uptake,
\item bespoke light threshold $\Rightarrow$ asymmetric response in POD,
\item temperature acclimation $\Rightarrow$ amplified effect of light threshold and drought.
\end{itemize}
The exact order of magnitude of these effects varies for individual species or PFT and between years, but the thus predicted ozone uptake is in any case larger than for MM default parameterizations. This indicates a considerable underestimation of ozone risk for subarctic species. Drought effects (SWP) do only matter if a bespoke GS is not taken into account.

In more detail for downy birch (Fig.~\ref{fig:pody_rel}a) all simulations exceed the CL by $5-25\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. Downy birch displays the largest range of ozone uptake among all species or PFTs ($10-31\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In particular, the sensitivity to the light threshold is especially pronounced, e.g. $\Delta\mathrm{POD_1} = ^{+2.5}_{-3.6}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2019. Without bespoke GS 2018 stands out as a higher ozone damage risk year compared to 2019. However, taking the later onset and overall shorter GS in 2018 into account, this is reversed for the boreal acclimation type. SWP induced reduction in ozone uptake is only effective in 2018 and without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold with respect to MM ranges between $(1.4-3.6)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.7-7.7)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to acclimation in temperature and therefor the least biased. Norway spruce displays a medium sensitivity to the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

For perennial grassland (Fig.~\ref{fig:pody_rel}c), all simulations with bespoke GS stay below the CL. Perennial grassland shows the smallest ozone uptake, but a similarly large bias as downy birch. We also find a reversal for predicted ozone damage risk in $\mathrm{POD_1}$ between 2018 and 2019 for the boreal type. Perennial grassland shows the lowest sensitivity to the light threshold and SWP is only relevant without bespoke GS. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(2.2-3.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(4.4-6.4)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019 with maximum systematic uncertainties due to light acclimation of $^{+0.3}_{-0.2}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+0.3}_{-0.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019). (a) Downy birch; (b) Norway spruce; (c) Perennial grassland. The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

The estimated biomass reduction in \unit{\%} for the bespoke temperature response function of type boreal in accordance to \citet{ICP:MappingManual2017, ESPR:Hayes2021} is shown in Table~\ref{tab:biomass_reduction}. The estimated uncertainty range corresponds to the deviation given by the $\pm 20\,\unit{\%}$ variations in light threshold and SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass are predicted for any of the two years. Norway spruce is least affected. The systematic uncertainties for downy birch are one order of magnitude larger than for Norway spruce and perennial grassland. For all species or PFTs the interannual variability is larger than the systematic uncertainties estimated. In comparison with average results for the default MM parameterization (with bespoke GS), we find a non negligible (above systematic uncertainties) low-bias for both downy birch and perennial grassland. For Norway spruce, the default MM parameterization average lies within the given systematic uncertainty range.

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} for bespoke temperature response function of type boreal with bespoke GS and SWP=off. Uncertainty ranges reported here correspond to deviations by the $\pm 20\,\unit{\%}$ variation in light threshold and SWP=on, respectively. For comparison results for MM parameterization are shown with standard deviations over both years and SWP treatments.}
  \label{tab:biomass_reduction}
\begin{tabular}{lllll}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
\middlehline
$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}
